---
title: "biostat776-project2-Shengyuan-Luo"
author: "davidluoshy"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Project instruction: <https://lcolladotor.github.io/jhustatcomputing/projects/project-2/>

# **Part 1: Fun with functions**

### **Part 1A: Exponential transformation**

The exponential of a number can be written as an infinite series expansion of the formexp⁡(x)=1+x+x22!+x33!+⋯

Of course, we cannot compute an infinite series by the end of this term and so we must truncate it at a certain point in the series. The truncated sum of terms represents an approximation to the true exponential, but the approximation may be usable.

Write a function that computes the exponential of a number using the truncated series expansion. The function should take two arguments:

-   `x`: the number to be exponentiated

-   `k`: the number of terms to be used in the series expansion beyond the constant 1. The value of `k` is always ≥1.

For example, if k=1, then the `Exp` function should return the number 1+x. If k=2, then you should return the number 1+x+x2/2!.

Include at least one example of output using your function.

```{r}
f <- function(x,k) {
if (k == 1) {
    return(1+x)
  } else if (k <1) {
    stop("Error: The number of terms to be used in the series expansion  cannot be <1")
  } else {
  output =x + 1
for (i in 2:k) {
  output =output +(x**i)/factorial(i)
}
    return(output)
  }
}
  
f(4,2)
```

### **Part 1B: Sample mean and sample standard deviation**

Next, write two functions called `sample_mean()` and `sample_sd()` that takes as input a vector of data of length N and calculates the sample average and sample standard deviation for the set of N observations.

x¯=1N∑i=1nxis=1N−1∑i=1N(xi−x―)2As a primer, the ∑i=13xi is equivalent to doing the sum of xi starting from i=1, increasing i by 1 each then, up to when i=3. That is, x1+x2+x3.

Include at least one example of output using your functions.

-   You can assume that the input value `x` will always be a *vector* of numbers of length *N*.

-   Do not use the `mean()` and `sd()` functions in R.

```{r}
sample_mean <- function(x) {
    print(sum(x)/length(x))
}

sample_sd <- function(x) {
  total <- 0
  for (i in seq_along(x)) {
   total = total + (x[i]-mean(x))^2
  }
  print(sqrt(total/(length(x)-1)))
}

sample_mean(c(1,8,5,3))
sample_sd(c(1,8,5,3))
```

### **Part 1C: Confidence intervals**

Next, write a function called `calculate_CI()` that:

1.  There should be two inputs to the `calculate_CI()`. First, it should take as input a vector of data of length N. Second, the function should also have a `conf` (=1−α) argument that allows the confidence interval to be adapted for different α.

2.  Calculates a confidence interval (CI) (e.g. a 95% CI) for the estimate of the mean in the population. If you are not familiar with confidence intervals, it is an interval that contains the population parameter with probability 1−α taking on this form

x¯±tα/2,N−1sx¯=(x¯−tα/2,N−1sx¯,x¯+tα/2,N−1sx¯)

where tα/2,N−1 is the value needed to generate an area of α/2 in each tail of the t-distribution with N−1 degrees of freedom and sx¯=sN is the standard error of the mean. For example, if we pick a 95% confidence interval and N=50, then you can calculate tα/2,N−1 as

```{r}
alpha <- 1 - 0.95
degrees_freedom <- 50 - 1
t_score <- qt(p = alpha / 2, df = degrees_freedom, lower.tail = FALSE)
```

3.  Returns a named vector of length 2, where the first value is the `lower_bound`, the second value is the `upper_bound`.

    ```{r}
    calculate_CI <- function(x, conf = 0.95) {
      alpha <- 1 - conf
      degrees_freedom <- length(x) - 1
      t_score <- qt(p = alpha / 2, df = degrees_freedom, lower.tail = FALSE)
      
      lower_bound <- mean(x) - t_score * sd(x)/sqrt(length(x)) 
      upper_bound <- mean(x) + t_score * sd(x)/sqrt(length(x)) 
      
      print(lower_bound)
      print(upper_bound)
    }
    calculate_CI(c(1,8,5,3), 0.8)
    calculate_CI(c(1,8,5,3), 0.95)
    ```

Include example of output from your function showing the output when using two different levels of `conf`.

# **Part 2: Wrangling data**

In this part, we will practice our wrangling skills with the tidyverse that we learned about in module 1.

### **Data**

The two datasets for this part of the assignment comes from [TidyTuesday](https://www.tidytuesday.com/). Specifically, we will use the following data from January 2020. To avoid re-downloading data, we will check to see if those files already exist using an `if()` statement:

```{r}
library(here)
file_url_temperature <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/temperature.csv"
local_file_path_temperature <- "/Users/s/Desktop/biostat776-project2-Shengyuan-Luo/data/temperature.csv"
download.file(file_url_temperature, destfile = local_file_path_temperature)

file_url_rainfall <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/rainfall.csv"
local_file_path_rainfall <- "/Users/s/Desktop/biostat776-project2-Shengyuan-Luo/data/rainfall.csv"
download.file(file_url_rainfall, destfile = local_file_path_rainfall)

library(readr)
rainfall <- read_csv(here("data", "rainfall.csv"))
temperature <- read_csv(here("data", "temperature.csv"))
```

Now we can look at the data with `glimpse()`

```{r}
library(tidyverse)
glimpse(rainfall)
```

If we look at the [TidyTuesday github repo](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020#2020-data) from 2020, we see this dataset contains temperature and rainfall data from Australia.

Here is a data dictionary for what all the column names mean:

-   <https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-01-07/readme.md#data-dictionary>

### **Tasks**

Using the `rainfall` and `temperature` data, perform the following steps and create a new data frame called `df`:

1.  Start with `rainfall` dataset and drop any rows with NAs.

    ```{r}
    rainfall_complete <- na.omit(rainfall)
    glimpse(rainfall_complete)
    ```

2.  Create a new column titled `date` that combines the columns `year`, `month`, `day` into one column separated by “-”. (e.g. “2020-01-01”). This column should not be a character, but should be recognized as a date. (**Hint**: check out the `ymd()` function in `lubridate` R package). You will also want to add a column that just keeps the `year`.

3.  Using the `city_name` column, convert the city names (character strings) to all upper case.

4.  Join this wrangled rainfall dataset with the `temperature` dataset such that it includes only observations that are in both data frames. (**Hint**: there are two keys that you will need to join the two datasets together). (**Hint**: If all has gone well thus far, you should have a dataset with 83,964 rows and 13 columns).

    ```{r}

    ```

# **R session information**

```{r}
options(width = 120)
sessioninfo::session_info()
```
